### 2013-05-31 doc/nips2013/appendix.tex

In the directed case, we assume a directed acyclic graph over $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Next, define a model as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To consolidate notation,
define $\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}. \aside{directed}
\end{align}

In this section, we describe parameter estimation in other models using our
framework.
a two-layer sigmoid belief network \cite{saul96sigmoid}.

@@

In \refsecs{threeViewMixtureModel}{generalModels}, we focused for simplicity on
undirected models, but we parameter estimation in directed models is also
possible.
In the directed mixture model, we have
\begin{align}
p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h),
\end{align}
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.

@@

\subsection{Factorial hidden Markov models}

\Fig{figures/factorialHMM}{0.7}{factorialHMM}{
Parameter in a factorial HMM.
}

\cite{ghahramani97fhmm}

>>>>

We have so far focused on undirected models for simplicity,
but our techniques also apply to directed models in which each
local conditional distribution is a log-linear model.

The setting is as follows:~Assume we have a directed acyclic graph over
nodes $V$,
and define $\pi(j) \subset V$ to be the parents of each node $j \in V$.
Define the joint distribution as a product of locally-normalized models:
\begin{align}
  p_\theta(z) = \prod_{j \in V} p_\theta(z_j \mid z_{\pi(j)}), \quad\quad
  p_\theta(z_j \mid z_{\pi(j)}) = \exp \{ \phi_j(z_j, z_{\pi(j)})^\top\theta - A_j(\theta; z_{\pi(j)}) \}.
\end{align}
Here, each local model has its own local features $\phi_j(z_j, z_{\pi(j)})$
and its own local log-partition function $A(\theta; z_{\pi(j)})$, which depends on values of its parent.
To consolidate notation with \refeqn{undirectedSetup},
define the sum of all the feature vectors
$\phi(z) = \sum_{j \in V} \phi(z_j, z_{\pi(j)})$
and $A(\theta; z) = \sum_{j \in V} A_j(\theta; z_{\pi(j)})$,
so that:
\begin{align}
  p_\theta(z) \eqdef \exp\{ \phi(z)^\top\theta - A(\theta; z) \}.
\end{align}

@@

In this section, we demonstrate parameter estimation for the directed mixture
model (\reffig{directedModels}(a)).  This is the usual mixture model, defined
as follows:
\begin{align}
p_\theta(x,h) = p_\theta(h) \prod_{j=1}^3 p_\theta(x_j \mid h),
\end{align}
where
$p_\theta(h) \eqdef \exp\{\phiprior(h)^\top\theta - \Aprior(\theta)\}$
and
$p_\theta(x_j \mid h) \eqdef \exp\{\philocal(x_j, h)^\top\theta - \Alocal(\theta; h)\}$.

@@

\Fig{figures/directedModels}{0.3}{directedModels}{
Various directed versions of the undirected models.
}

\Fig{figures/factorialModels}{0.3}{factorialModels}{
Directed factorial models.
}

\subsection{Other models}

\reffig{directedModels} shows examples of the directed models
for which we have already derived undirected estimates.
In general, the aggregation
and factorization steps (Steps 1 and 2) in \refsec{threeViewMixtureModel}
remain the same as the undirected cases since these steps
only rely on the conditional independence structure of the models,
not particular parametrizations.

\subsection{Factorial models}

For factorial models, the directed and undirected version do differ in their
conditional independence statements.
For example, consider the 
the directed factorial mixture model,
also known as a sigmoid belief network \citep{saul96sigmoid},
in \reffig{factorialModels}(a).
Here ,$h_1$ and $h_2$ are marginally independent in the directed
case but conditionally indepdent in the undirected case.
However, since our algorithm treats $(h_1,h_2)$ as a composite variable $c$
anyway, it is not sensitive to these differences.
The factorial HMM \cite{ghahramani97fhmm} is similar and can be handled
similar to our treatment of HMMs plus unshuffling factorization.

### 2013-05-31 doc/nips2013/factorial.tex

  \item To do this last step, construct a linear program with $\nh^s \cdot d$ constraints, one for each entry of $B$.
  It will have $s\nh \cdot d + \nh^s$ variables,
  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constants $Z_\ba$.
  Note that $Z_\ba$ is actually a determinstic function of the the vectors, but since it is a non-linear one,
  it is difficult to work with; therefore, we create a separate variable to keep the problem linear.

>>>>

  \item To do this last step, construct a linear program with $\nh^s \cdot d$ constraints, one for each entry of $B$.
  It will have $s\nh \cdot d + \nh^s$ variables,
  one for each of $s\nh$ vectors $v_{ia}$ and $\nh^s$ for the normalization constants $Z_\ba$.
  Note that $Z_\ba$ is actually a deterministic function of the the vectors, but since it is a non-linear one,
  it is difficult to work with; therefore, we create a separate variable to keep the problem linear.
